A new method to estimate dark matter halo concentrations

### 2014-11-05 paper/article.tex
first pass intro

\title[Bayesian halo concentrations]{Dark matter halo
  concentrations with a bayesian approach}
\author[Poveda \& Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright
  Christian Poveda$^{1}$ \&
  Jaime E. Forero-Romero$^{1}$
}
\vspace*{6pt}\\
$^{1}$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
}
\maketitle

@@

\section{Introduction}
\label{sec:introduction}
In the context of the concordance cosmology, the matter contents of
the Universe are dominated by an unknown dark matter which behaves as
a collisionless fluid only perturbed by gravity.

Numerical experiments are able to follow the non-linear evolution of
dark matter down to galactic scales. On these scales a common
approximation considers the matter distribution as spherical with a
radial depedence. The function describing this distribution is known
as a radial density profile.

The consensus is that density profiles are almost universal;
almost independet of cosmological parameters and similar for different
objects after some simple physical re-scalig is applied.

The detailed halo structure influences the hosted galaxies.

The concentration parameter is imporant in galaxy formation studies.

Different studies vary about the result of the concentration-mass
relationship.

We propose a new way to obtain the concentration for halos in N-body
simulations.

We use the integrated mass profile because it
allows us to use the data directly with the simulation without binning
the particle positions and estimating a density.

As an application we study the mass concentration relationship.

\citep{NFW}

@@

\section{Halo Samples}

@@

\subsection{Mock Halos}
\label{sec:results_mocks}

\subsection{Simulation Data}
\label{sec:results_mocks}

@@

\subsection{Comparison against other methods}
We compared this method against two methods: The first one consists in using shells for estimating the density in function of the radius and using the same MCMC method for fitting and the second one consists in using the circular velocity $V(r)=\sqrt{GM(<r)/r}$ and the relation for the NFW profile
\begin{equation}
\frac{V_{max}}{V(r_{v})} = \sqrt{\frac{0.216c}{M(r_{v};c)}}
\end{equation}
Where $V_{max}$ is the maximum velocity, to find the value of the concentration. We tested the three methods using the same data from the Mock Halos test. In order to see the accuracy of each method depending on the number of particles we define

@@

\subsection{Concentration as a function of halo mass}
Aditionally we compared those three methods plotting the concentration as a function of the number of particles of each halo

@@

The bold lines in Figure \ref{fig:concentrations} corresponds to the median and the thinner lines corresponds to the quartiles. Showing that our method has values ​​ranging from those obtained from the other two methods.Also we can see that the concentration obtained by the velocity method is higher than that obtained by the density method, this fact is consistent with FIXME.
\subsection{Implication for comparisons against observations}
FIXME: Ask to Jaime

\section{Conclusions}
\label{sec:conclusions}
FIXME: Ask to Jaime

@@

\end{document}

>>>>

\title[New approach to Halo concentrations]{A new method to estimate
  dark matter halo concentrations}
\author[Poveda \& Forero-Romero]{
\parbox[t]{\textwidth}{\raggedright
  Christian Poveda$^{1}$ \&
  Jaime E. Forero-Romero$^{1}$
}
\vspace*{6pt}\\
$^{1}$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
}
\maketitle

@@

\section{Introduction}
\label{sec:introduction}
In the concordance cosmology paradigm the the matter content of the
Universe are dominated by dark matter which behaves as a collisionless
fluid under the influence of gravity. In the last three decades
numerical experiments have made possible the simulation of large
volumes of a Universe dominated by dark matter.

One of the most striking results of these simulations is that dark
matter clumps on galactic scales follows a universal density profile
which in a first approximation is spherical symmetric and only
dependent on the radial coordinate. These profiles seem to be
universal; independent of the cosmological parameters and self-similar
for different spatial scapes after an addecuate re-scaling is
appplied.

One the most commong parameterization of this density is known as the
Navarro-Frenk-White (NFW) profile \citep{NFW}. This profile is a
double power law in radius, where the transition between the two
happens at the so-called scale radius $r_s$. The ratio between the
scale radius and the virial radius of the halo $R_v$ is known as the
concentration $c=R_V/r_s$, a quantity that has been found to be
correlated with the halo mass.

The relationship between halo mass and concentration is in principle
accesible to observations and provides a potential test of LCDM on
galactic scales. With this promise a great deal of effort has been
invested in calibrating this relationship with simulations but also
finding the best possible way to constraint it with observations.

From the computational point of view there are two main methods to
estimate the concetration parameter of a dark matter halo in a N-body
simulation. The first method takes the particles comprosing the halos
and bins them in logarithmic radii to estimate the density in each
bin, then it proceeds with a fit of this density estimation as a
function of radius. A second method uses an analytical property of the
NFW that related the maximum of the ratio of the circular velocity to
the virial velocity. In this method at each particle radius this ratio
is estimated to find the maximum value for this ratio; this value is
used to find the roots of an equation which represent the
concentration parameter.

The first method although is straightforward to apply has two main
disadvantages. First of all it requires a large number of particles in
order to have a proper density estimate in each bin. For a very low
number of particles is hard to define a large number of bins to
proceed with the fit. The second problem is that, as in any process
involving data binning, there is not an way to estimate the optimal
bin size, which in turn can affect the results of the fit.

The second method solves the two problems mentioned above. It works
with low particles numbers and does not involve data binning. However,
it effectively takes into account a single data point and discards the
behaviour of the ratio $V_{cirv}/V_{vir}$ below and above its
maxima.

In this paper we propose a new method to estimate the dark matter halo
concentration from N-body simulation results. This method has two
advantages with respect to the methods mentioned above. First, it does
not involve data binning. Second, it does not throw away data
points. Third, it allows for an straightforward estimation of the
uncertainties in the concentration parameter.

Our method consists in building the cumulative mass profile from the
particle data in the N-body simulation to find the best possible
concentration value using a Markov Chain Monte Carlo methodology by
comparing the data against the analytical expectation.

This paper is structured as follows. In Section \ref{sec:basics} we
review the basic properties of the NFW density profile and define the
basic notation for the rest of the paper. Next in Section
\ref{sec:method} we present our new method to estimate the halo
concentration, payin special attention to the bayesian framework to
find the most probable value and its uncertaintt. In Section
\ref{sec:halo_sample} we demonstrate the power of our method with two
different halo samples; the first one generated under controlled
conditions and the second taken from a cosmological N-body
simulation. Next in \ref{sec:discussion} we discuss these results by
comparing them against other methods to estimate the concentration and
comment on the consequences for the concentration-mass
relationship. We present our conclusions in Section \ref{sec:conclusions}.

@@

@@

@@

\subsection{Comparison against other methods}
\label{sec:comparison}
We compared this method against two methods: The first one consists in
using shells for estimating the density in function of the radius and
using the same MCMC method for fitting and the second one consists in
using the circular velocity $V(r)=\sqrt{GM(<r)/r}$ and the relation
for the NFW profile 
\begin{equation}
\frac{V_{max}}{V(r_{v})} = \sqrt{\frac{0.216c}{M(r_{v};c)}}
\end{equation}
Where $V_{max}$ is the maximum velocity, to find the value of the concentration. We tested the three methods using the same data from the Mock Halos test. In order to see the accuracy of each method depending on the number of particles we define

@@

\subsection{Impact on the mass-concentration relationship}
Aditionally we compared those three methods plotting the concentration as a function of the number of particles of each halo

@@

The bold lines in Figure \ref{fig:concentrations} corresponds to the median and the thinner lines corresponds to the quartiles. Showing that our method has values ​​ranging from those obtained from the other two methods.Also we can see that the concentration obtained by the velocity method is higher than that obtained by the density method, this fact is consistent with FIXME.
\subsection{Implication for comparisons against observations}
FIXME: Ask Jaime

\section{Conclusions}
\label{sec:conclusions}
FIXME: Ask Jaime

@@

\verb"http://adsabs.harvard.edu/abs/2014ApJ...795..163U"
\verb"http://adsabs.harvard.edu/abs/2012ApJS..199...25P" Figura 3!
\end{document}
